WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2019-03-06 review/referee_response_0.txt

REFEREE COMMENT

2) As mentioned in 1), another possible interpretation for explaining the
deviation from a linear ephemeris of the TESS mid-transit times could be the
presence of anomalies in the light curves due to stellar activity. It has been
well established that WASP-4 is an active star with a rotation period (P) in
between 20-40 days (P ~ 22 days and P ~ 34 days were determined by
Sanchis-Ojeda et al. 2011 and Hoyer et al. 2013, respectively), which is
expected for a main-sequence G7 star. Given that the TESS mid-transits span
around 30 days, wouldn't it be possible that these measurements be affected by
the existence of unseen stellar spots? Regarding this point, I would like to
see an extra paragraph in the Section "Interpretation" with a discussion about
this scenario.

  > RESPONSE: We have considered this possibility, and have now added
  > the following text in Section 3.3:

  > To explore the effect of possible spot-crossing events on the TESS
  > transit time measurements, we performed a separate test.  We
  > injected triangular spot-anomalies with amplitude $0.03\%$ and
  > duration 30 minutes at random phases into each transit.  The
  > amplitudes were larger than the spot-crossing anomalies observed
  > by \citet{southworth_high-precision_2009} and
  > \citet{sanchis-ojeda_starspots_2011}, and the durations were
  > comparable.  Spots of these amplitudes resemble the anomalies
  > present in transits ``1360.54'' and ``1372.58'' of Figure
  > \ref{fig:lightcurves}: they are visible, but at low confidence.
  > 
  > With spots injected, we repeated our measurement of the transit
  > times.  On average, the measured transit times did not change
  > after injecting spots, because the flux deviations are equally
  > likely to occur in the first and second halves of the transit.
  > For individual transits, there were no cases for which the timing
  > deviation was larger than one minute.  The largest shifts occur
  > when the spot anomaly occurs during transit ingress or egress, in
  > which case the measured mid-time is shifted either late or early
  > by between 30 and 50 seconds \citep[qualitatively similar to
  > results found by][]{ioannidis_how_2016}.
  > 
  > Therefore the TESS observations could hypothetically all be skewed
  > early if there were spot-crossing events during every egress.  Two
  > arguments rule out this possibility.  (1) The lightcurve residuals
  > do not show evidence for these events.  (2) The stellar rotation
  > period is between 20 and 40 days, and the sky-projected stellar
  > obliquity is less than 10 degrees
  > \citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011,hoyer_tramos_2013}.
  > Since the planet orbits every 1.3 days, requiring that spot
  > anomalies always occur during egress would be equivalent to
  > requiring a stellar spot distribution that is exquisitely (and
  > thus implausibly) distributed to match the planet egress times.

@@

REFEREE COMMENT

vii) Given that the mid-transit times values used in this work are taken from
different authors with different methodologies to perform the fitting and
detrending of the light curves, could the authors mention how this could affect
their results?

  > RESPONSE: We have added the following paragraphs at the end of
  > Section 3.3:

  > There is a final concern that is difficult to address.  We collected
  > the mid-transit time values derived by different authors, who used
  > heterogeneous methods to fit and detrend their lightcurves.  We have
  > also assumed that these authors have correctly documented the time
  > systems in which the data are reported.  Further, though many choices
  > in transit-fitting ({\it e.g.}, parametrization of limb-darkening and
  > eccentricity) do not affect transit mid-time measurements, different
  > detrending approaches can asymmetrically warp transits and shift
  > mid-transit times.  The magnitude of this systematic effect is hard to
  > quantify, but the situation is fairly clear from
  > Figure~\ref{fig:times}.  Many independent authors provided transit
  > measurements shortly after WASP-4b's discovery, and the data are
  > consistent with each other. \citet{huitson_gemini_2017} provided the
  > most important data from epochs 0-1000. If their data were
  > systematically affected by detrending choices or time-system confusion
  > at the level of several times their reported uncertainties, then it
  > possible that the orbital period is constant despite the evidence in
  > the TESS data.  For this reason, we paid careful attention to the
  > \citet{huitson_gemini_2017} data set, and corresponded with the
  > authors to confirm that their results are not affected by systematic
  > effects of the required amplitude.
  > 
  > None of the concerns mentioned in this subsection seem likely to
  > explain the observed timing variations. We proceed by considering
  > possible astrophysical explanations.

### 2019-03-06 paper/ms.tex
arxiv submission ready

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  \replaced{The residuals are the observed times minus the calculated
  times assuming a constant period. The darkest points correspond to
  the most precise data.}{The vertical axis shows the observed times
  minus the calculated times assuming a constant period for transits
  ({\it top}) and occultations ({\it bottom}). Darker points
  correspond to more precise data.} The constant-period model (gray
  line) is a poor description of the data.  Models with a uniformly
  decreasing period (blue) or an eccentric, precessing orbit (orange)
  provide	better fits\deleted{ to the data}.  The red square
  represents the combination of all the TESS data and is for display
  purposes only.  The models were fitted to all of the individual
  transit times.
  \label{fig:times}
	}
\end{figure*}

@@

First, we collected the usable RV measurements from CORALIE, HARPS,
and HIRES.  We included the CORALIE measurements from
\citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010}, using
the homogeneous radial velocities calculated by the latter authors. We
included the HARPS values reported by \citet{pont_determining_2011},
which are identical to those from \citet{husnoo_observational_2012}.
We omitted the HARPS data points taken over three nights by
\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
because they were calculated using a different pipeline than the
longer-baseline \citealt{pont_determining_2011} HARPS measurements,
and necessary inclusion of an extra offset term would nullify their
statistical value.  Finally, we included the five HIRES measurements
taken over many years by \citet{knutson_friends_2014}.

@@

To summarize, only about one tenth of the observed period decrease can
be explained through a constant acceleration of the WASP-4 system's
center of mass.  However, given the limited amount and uneven time
coverage of the existing radial-velocity data (Figure~\ref{fig:rvs}),
it remains possible that the center of mass has a more complex motion,
perhaps due to a companion on an eccentric orbit \citep[{\it e.g.},
WASP-53 or WASP-81][]{triaud_peculiar_2017}.  It would be useful to
gather more radial-velocity data to confirm or refute this
possibility.

@@

For each planet, we asked: what fraction of the posterior for $t_{\rm
offset}$ is consistent with an offset worse than $81.6$ seconds?  For
WASP-4b, the answer is unsurprisingly 50\%.  For WASP-6b, the most
constraining object, about 1 sample in 2 million is consistent with
such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 103 samples
would be consistent with this timing offset ($2.3\sigma$), and in
WASP-46b, the limit is 1 in 49 samples ($2.0\sigma$).  For WASP-5b,
the predicted time is too imprecise to rule out timing offsets at the
necessary amplitude.  Multiplying the three independent probabilities
for WASP-5b, 6b, and 18b, we can rule out $t_{\rm offset} < -81.6\
{\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.

>>>>

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  \replaced{The residuals are the observed times minus the calculated
  times assuming a constant period. The darkest points correspond to
  the most precise data.}{The vertical axis shows the observed times
  minus the calculated times assuming a constant period for transits
  ({\it top}) and occultations ({\it bottom}). In the upper panel,
  darker points correspond to more precise data.} The constant-period
  model (gray line) is a poor description of the data.  Models with a
  decreasing period (blue) or an eccentric, precessing orbit (orange)
  provide	better fits\deleted{ to the data}.  The red square
  represents the combination of all the TESS data and is for display
  purposes only.  The models were fitted to all of the individual
  transit times.
  \label{fig:times}
	}
\end{figure*}

@@

First, we collected the usable RV measurements from CORALIE, HARPS,
and HIRES.  We included the CORALIE measurements from
\citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010}, using
the homogeneous radial velocities calculated by the latter authors. We
included the HARPS values reported by \citet{pont_determining_2011},
which are identical to those from \citet{husnoo_observational_2012}.
We omitted the HARPS data points taken over three nights by
\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
because they were calculated using a different pipeline than the
longer-baseline \citeauthor{pont_determining_2011} HARPS measurements,
and necessary inclusion of an extra offset term would nullify their
statistical value.  Finally, we included the five HIRES measurements
taken over many years by \citet{knutson_friends_2014}.

@@

To summarize, only about one tenth of the observed period decrease can
be explained through a constant acceleration of the WASP-4 system's
center of mass.  However, given the limited amount and uneven time
coverage of the existing radial-velocity data (Figure~\ref{fig:rvs}),
it remains possible that the center of mass has a more complex motion,
perhaps due to a companion on an eccentric orbit \citep[{\it e.g.},
WASP-53 or WASP-81,][]{triaud_peculiar_2017}.  It would be useful to
gather more radial-velocity data to confirm or refute this
possibility.

@@

For each planet, we asked: what fraction of the posterior for $t_{\rm
offset}$ is consistent with an offset worse than $81.6$ seconds?  For
WASP-4b, the answer is unsurprisingly 50\%.  For WASP-6b, the most
constraining object, about 1 sample in 2 million is consistent with
such a timing offset ($4.9\sigma$).  For WASP-18b, 1 in 103 samples
would be consistent with this timing offset ($2.3\sigma$), and in
WASP-46b, the limit is 1 in 49 samples ($2.0\sigma$).  For WASP-5b,
the predicted time is too imprecise to rule out timing offsets at the
necessary amplitude.  Multiplying the three independent probabilities
for WASP-6b, 18b, and 46b, we can rule out $t_{\rm offset} < -81.6\
{\rm seconds}$ at $6.4\sigma$, or about about 1 part in 11 billion.
